---
title: 'Data Analysis Assignment #1 (50 points total)'
author: "Strouse, Logan"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include = FALSE}
# DO NOT ADD OR REVISE CODE HERE
knitr::opts_chunk$set(echo = FALSE)

```

-----

Submit both the .Rmd and .html files for grading. You may remove the instructions and example problem above, but do not remove the YAML metadata block or the first, "setup" code chunk.  Address the steps that appear below and answer all the questions. Be sure to address each question with code and comments as needed.  You may use either base R functions or ggplot2 for the visualizations.

-----

The following code chunk will (a) load the ggplot2 and gridExtra packages, assuming each has been installed on your machine, (b) read-in the abalones dataset, defining a new data frame, "mydata," (c) return the structure of that data frame, and (d) calculate new variables, VOLUME and RATIO. If either package has not been installed, you must do so first via *install.packages()*; e.g. *install.packages("ggplot2")*. You will also need to download the abalones.csv from the course site to a known location on your machine.

```{r analysis_setup1, eval = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# a) Load the ggplot2 and gridExtra packages.

library(ggplot2)
library(gridExtra)
library(knitr)
library(plyr)
library(dplyr)

# b) Use read.csv() to read the abalones.csv into R, assigning the data frame to "mydata."

# mydata <- read.csv(file.path("c:/Rabalone_new/", "abalones.csv"), sep = ",")
mydata <- read.csv(file.path("/Users/lstrouse/Desktop/MSDS401/abalones.csv"), sep = ",")

# c) Use the str() function to verify the structure of "mydata." You should have 1036 observations
# of eight variables.

str(mydata)

# d) Define two new variables, VOLUME and RATIO. Use the following statements to define VOLUME and
# RATIO as variables appended to the data frame "mydata."

mydata$VOLUME <- mydata$LENGTH * mydata$DIAM * mydata$HEIGHT
mydata$RATIO <- mydata$SHUCK / mydata$VOLUME

```

-----

(1)(a) (1 point) Use *summary()* to obtain and present descriptive statistics from mydata.  

```{r Part_1a, eval = TRUE, echo = FALSE}
summary(mydata)
```

**Question (1 point):  Briefly discuss the variable types and distributional implications such as potential skewness and outliers.**

***Answer: There are a total of 1036 observations and 10 variables so far. Based on looking at these variables you can see some potential outliers and potential skewness. There are also different types of variables.Some of the variables are quantitative and some qualitative. SEX would be a qualitative variables, since there is a finite 3 types it could be. LENGTH, DIAM, HEIGHT, WHOLE, SHUCK are a few examples of the continuous variables. We also have Ratio level data, which could be HEIGHT, VOLUME or LENGTH variables. SHUCK,WHOLE and VOLUME are the variables that concern me most based on the summary tables for outliers. This is due to how big the range appears to be in relation to the mean.   ***

(1)(b) (1 point) Generate a table of counts using SEX and CLASS. Add margins to this table (Hint: There should be 15 cells in this table plus the marginal totals. Apply *table()* first, then pass the table object to *addmargins()* (Kabacoff Section 7.2 pages 144-147)).  Lastly, present a barplot of these data.  

```{r Part_1b, eval = TRUE, echo = FALSE}
mytable<- with(mydata,table(SEX,CLASS))
addmargins(mytable)
barplot(mytable,beside = TRUE,ylab = "Frequency",xlab = "Class", col = c("red","blue","green"),main = "CLASS membership, SEX-differentiated",legend = row.names(mytable))
```

**Question (1 point):  Discuss the sex distribution of abalones.  What stands out about the distribution of abalones by CLASS?**

***Answer: I found this graph to be very insighful. The sex distribution makes sense. The smallest abalones should be the A1 class since they are the youngest. Eventually, towards the middle tiers you see them starting to tapper off into smaller poplutions of the class. Looking at the chart that was created as well with the sums, you can see that the sex distribuion of the abalones is relatively even. There is a slight skew towards there being sightly more Males than Females or Infants.I would be curious to see out of that group of 329 infants, how many end up becoming male or female? Is there more of a trend towards more females being created or are males just living longer or not being harvested as frequently? This is particularly curious in the A1-A3 classes where the males significantly outnumber the females.  ***

(1)(c) (1 point) Select a simple random sample of 200 observations from "mydata" and identify this sample as "work". Use *set.seed(123)* prior to drawing this sample. Do not change the number 123. (If you must draw another sample from mydata, it is imperative that you start with *set.seed(123)*, otherwise your second sample will not duplicate your first sample or the "work" sample used for grading your report.) (Kabacoff Section 4.10.5 page 87) 

Using this sample, construct a scatterplot matrix of variables 2-6 with *plot(work[, 2:6])* (these are the continuous variables excluding VOLUME and RATIO). The sample "work" will not be used in the remainder of the assignment. 

```{r Part_1c, eval = TRUE, echo = FALSE}
set.seed(123)
work <- mydata[sample(nrow(mydata),200,replace=FALSE),]
plot(work[, 2:6])
```

-----

(2)(a) (1 point) Use "mydata" to plot WHOLE versus VOLUME. Color code data points by CLASS.

```{r Part_2a, eval = TRUE, echo = FALSE}
ggplot(mydata, aes(VOLUME,WHOLE,color=CLASS)) + geom_point() 
```

**Question (2 points):  What does the wedge-shaped scatter of data points suggest about the relationship between WHOLE and VOLUME?  Interpret this plot taking into account abalone physical measurements of length, diameter and height and the displays shown in (1)(c).**

***Answer: There is a correlation and/or a strong relationship and between WHOLE and VOLUME based upon the scatterplot. It is logical for the volume of an abalone to increase with the whole weight of the same abalone. As an abalone grows and gets bigger the volume (length*diameter*height) should reflect that. One thing I do find interesting about the plot was the amount of A3 abalones in the upper right quadrant. I would have expected them to be moe centrally located in the plot and not with the A4/A5 Classes. I would have to do some analysis on those points to see if they are just outliers. In regards to the sample on 1C, it appears that HEIGHT and SHUCK are some of the least correlated variables, due to how they are dispersed on the chart.     ***

(2)(b) (2 points) Use "mydata" to plot SHUCK versus WHOLE with WHOLE on the horizontal axis. Use a different color for each age class.  As an aid to interpretation, determine the maximum value of the ratio of SHUCK to WHOLE.  Add to the chart a straight line with zero intercept using this maximum value as the slope of the line. If you are using the 'base R' *plot()* function, you may use *abline()* to add this line to the plot. Use *help(abline)* in R to determine the coding for the slope and intercept arguments in the functions. If you are using ggplot2 for visualizations, *geom_abline()* should be used.

```{r Part_2b, eval = TRUE, echo = FALSE}
ggplot(mydata, aes(WHOLE,SHUCK,color=CLASS)) + geom_point() + geom_abline(slope = max(mydata$SHUCK/mydata$WHOLE),intercept = 0,mapping = FALSE,Col ="green")
```

**Question (2 points):  How does the variability in this plot differ from the plot in (a)?  Compare the two displays.  Keep in mind that SHUCK is a part of WHOLE.**

***Answer: The data points are more tightly bound together at 0 until you get closer to the max of the plot. The slope of the line does not follow the mean of the data since it picks the max ratio to follow. What is concerning to me is that as you get the bigger whole abalones, it does not appear that the shuck yield is as good. Some of the biggest shuck yields apprear to be coming from the middle class (A3) abalones. Essentially, the second plot is telling us that the oldest abalones don't always produce the biggest shuck.   ***

-----

(3)(a) (2 points) Use "mydata" to create a multi-figured plot with histograms, boxplots and Q-Q plots of RATIO differentiated by sex. This can be done using *par(mfrow = c(3,3))* and base R or *grid.arrange()* and ggplot2. The first row would show the histograms, the second row the boxplots and the third row the Q-Q plots. Be sure these displays are legible.  

```{r Part_3a, eval = TRUE, echo = FALSE}
sex_label <- c("Female", "Infant", "Male")
names(sex_label) <- levels(mydata$SEX)
ggplot(mydata, aes(RATIO)) + geom_histogram(fill='blue') +
facet_grid(~ SEX, labeller = labeller(SEX = sex_label))
ggplot(mydata, aes(x = SEX, y = RATIO, group = SEX)) +
geom_boxplot(fill = "blue") + 
facet_grid(~ SEX, labeller = labeller(SEX = sex_label),scales="free")
slopes <- numeric()
for(i in levels(mydata$SEX)) {
slopes[i] <- diff(quantile(mydata[mydata$SEX == i, "RATIO"],
probs = c(0.25, 0.75))) /
diff(qnorm(c(0.25, 0.75)))
}
ints <- numeric()
for(i in levels(mydata$SEX)){
ints[i] <- quantile(mydata[mydata$SEX == i, "RATIO"],
probs = c(0.25)) - slopes[i] *
qnorm(c(0.25))
}
ab_df <- data.frame(SEX = as.character(levels(mydata$SEX)),
a = slopes,
b = ints)
ggplot(mydata, aes(sample = RATIO)) + geom_qq() + 
geom_abline(data = ab_df, aes(slope = a, intercept = b)) +
facet_grid(~ SEX, labeller = labeller(SEX = sex_label))

```

**Question (2 points): Compare the displays.  How do the distributions compare to normality? Take into account the criteria discussed in the sync sessions.**

***Answer: Based on the lecture from sync session three, I was able to determine that the distribution was normal and symetric with few outliers. The sample size was larger than the 30 or 40 min. that was suggested to use our methods. The QQ plots show that the data follows the line relatively well with the exception of Infants. There appears to be more variability with that sex, than with Male or Female. The boxplots do a good job of showing where the majority of the data points fall with those few outliers.    ***

(3)(b) (2 points) Use the boxplots to identify RATIO outliers. Present the abalones with these outlying RATIO values along with their associated variables in "mydata." Hint:  Construct a listing of the observations using the kable() function.  

```{r Part_3b, eval = TRUE, echo = FALSE}
female_ratio <- boxplot.stats(mydata[mydata$SEX == "F", "RATIO"])$out
infant_ratio <- boxplot.stats(mydata[mydata$SEX == "I", "RATIO"])$out
male_ratio <- boxplot.stats(mydata[mydata$SEX == "M", "RATIO"])$out
combined_ratio <- c(female_ratio, infant_ratio, male_ratio)
mydata[mydata$RATIO %in% combined_ratio, ]
kable(mydata[mydata$RATIO %in% combined_ratio, ])
```

**Question (2 points):  What are your observations regarding the results in (3)(b)?  **

***Answer: 3A contains the outliers for the problem. I find it interesting that it isn't just one sex producing the outliers. It appears to be evenly coming from each sex. One outlier theme that I think would be worth investigating would be the RINGS and CLASS variables. It appears that the outliers for the data frame are coming from the A1 and A3 classes. There are different explanations to these observations. I would have to do further statistical research to explain the meaning. Could it be that the rings were miss-counted or some other possible human error?  ***

-----

(4)(a) (3 points) With "mydata," display two separate sets of side-by-side boxplots for VOLUME and WHOLE differentiated by CLASS (Davies Section 14.3.2). Show five boxplots for VOLUME in one display and five boxplots for WHOLE (making two separate displays). Also, create two separate scatterplots of VOLUME and WHOLE versus RINGS. Present these displays in one graphic, the boxplots in one row and the scatterplots in a second row. Base R or ggplot2 may be used.


```{r Part_4a, eval = TRUE, echo = FALSE}
grid.arrange(
ggplot(mydata, aes(x = CLASS, y = VOLUME, group = CLASS)) +
geom_boxplot(fill = "blue"),
ggplot(mydata, aes(x = CLASS, y = WHOLE, group = CLASS)) +
geom_boxplot(fill = "blue"),
ggplot(mydata, aes(x = RINGS, y = VOLUME, group = CLASS)) +
geom_point(),
ggplot(mydata, aes(x = RINGS, y = WHOLE, group = CLASS)) +
geom_point(),
nrow = 2)
```

**Question (5 points) How well do you think these variables would perform as predictors of age?**

***Answer: I found these two sets of charts to be very insightful. Particularly the Scatter plots. The individual points clearly outline the extreme variability in the rings. There is a general trend towards larger class abalones as the VOLUME AND WHOLE  increase. A problem really makes itself visable on the A4 and A5 classes of the box plots. They look almost identical, which would mean that it is extrememly difficult to discern the age of an older abalone between A4 and A5, based on just VOLUME AND WHOLE. The boxplots also show in both charts a significant amount of outliers from A1 to A3. In recap, I think these variables are very difficult to predict age.   ***

-----

(5)(a) (2 points) Use *aggregate()* with "mydata" to compute the mean values of VOLUME, SHUCK and RATIO for each combination of SEX and CLASS. Then, using *matrix()*, create matrices of the mean values. Using the "dimnames" argument within *matrix()* or the *rownames()* and *colnames()* functions on the matrices, label the rows by SEX and columns by CLASS. Present the three matrices (Kabacoff Section 5.6.2, p. 110-111). You do not need to be concerned with the number of digits presented.

```{r Part_5a, eval = TRUE, echo = FALSE}
list('Volume',matrix(aggregate(VOLUME~CLASS + SEX,mydata, mean)$VOLUME, ncol = 5, byrow = 3,
dimnames = list(unique(aggregate(VOLUME~CLASS + SEX,mydata, mean)$SEX), unique(aggregate(VOLUME~CLASS + SEX,mydata, mean)$CLASS))))
list('Shuck',matrix(aggregate(SHUCK~CLASS + SEX,mydata, mean)$SHUCK, ncol = 5, byrow = 3,
dimnames = list(unique(aggregate(SHUCK~CLASS + SEX,mydata, mean)$SEX), unique(aggregate(SHUCK~CLASS + SEX,mydata, mean)$CLASS))))
list('Ratio',matrix(aggregate(RATIO~CLASS + SEX,mydata, mean)$RATIO, ncol = 5, byrow = 3,
dimnames = list(unique(aggregate(RATIO~CLASS + SEX,mydata, mean)$SEX), unique(aggregate(RATIO~CLASS + SEX,mydata, mean)$CLASS))))
```

(5)(b) (3 points) Present three graphs. Each graph should be generated with three separate lines appearing, one for each sex. The first should show mean RATIO versus CLASS; the second, average VOLUME versus CLASS; the third, SHUCK versus CLASS. This may be done with the 'base R' *interaction.plot()* function or with ggplot2.


```{r Part_5b, eval = TRUE, echo = FALSE, fig.width = 9}
ggplot(aggregate(RATIO~CLASS + SEX,mydata, mean)
, aes(CLASS,RATIO, group = SEX, col = SEX)) + geom_line() + geom_point() + ggtitle("MEAN RATIO PER CLASS")
ggplot(aggregate(VOLUME~CLASS + SEX,mydata, mean) 
, aes(CLASS,VOLUME, group = SEX, col = SEX)) + geom_line() + geom_point() + ggtitle("MEAN VOLUME PER CLASS")
ggplot(aggregate(SHUCK~CLASS + SEX,mydata, mean)
, aes(CLASS,SHUCK, group = SEX, col = SEX)) + geom_line() + geom_point() + ggtitle("MEAN SHUCK WEIGHT PER CLASS")

```

**Question (2 points):  What questions do these plots raise? Discuss.**

***Answer: I truely believe that these plots prove some of my earlier points. There are still some basic understandings happening that I don't want to loose sight of. As abalones age, they do get larger and their volume increases. With that, you would think the shuck would increase as well, but it doesn't. I think there are multiple reasons for this, but I can't empirically prove them in this data assignment. As shown by the mean shuck weight per class, there is a degradation in the shuck as the abalone ages in class from A4 to A5. I found that to be particularly interesting. What was that happeneing?  The mean volume per class also shows a very similar pattern as well. These charts added to my questions about the study and I would really like to see what is causing that. I have a feeling there are other variables, bysides the ones studied that we should be looking at.    ***

5(c) (3 points) Present four different boxplot displays using *par(mfrow = c(2, 2)*. The first line would show VOLUME by RINGS for the infants and the adults (factor levels "M" and "F" combined), The second line would show WHOLE by RINGS for the infants and the adults (factor levels "M" and "F" combined). Since the data are sparse beyond 15 rings, limit the displays to less than 16 rings.  Use ylim = c(0,1100) for VOLUME and ylim = c(0, 400) for WHOLE.  If you wish to reorder the displays for presentation purposes or use ggplot2 go ahead.

```{r Part_5c, eval = TRUE, echo = FALSE}
infants <- mydata[mydata$SEX == "I" & mydata$RINGS < 16, ]
adults <- mydata[mydata$SEX != "I" & mydata$RINGS < 16, ]

par(mfrow = c(1, 2))
boxplot(infants$VOLUME ~ infants$RINGS, main = "Infant Volume Rings", col = "green", xlab = "Rings", ylab = "Volume", ylim = c(0, 1100))
boxplot(adults$VOLUME ~ adults$RINGS, main = "Adult Volume Rings", col = "red", xlab = "Rings", ylab = "Volume", ylim = c(0, 1100))
boxplot(infants$WHOLE ~ infants$RINGS, main = "Infant Whole Weight Rings", col = "green", xlab = "Rings", ylab = "Whole Weight", ylim = c(0, 400))
boxplot(adults$WHOLE ~ adults$RINGS, main = "Adult Whole Weight Rings", col = "red", xlab = "Rings", ylab = "Whole Weight", ylim = c(0, 400))
par(mfrow = c(1, 1))
```

**Question (2 points):  Abalone growth is said to decline when they have more than ten rings.  Do you see trends in these plots to support this statement?**

***Answer: Based on the data of the 4 charts above, I agree with the statement that after 10 rings the abalone growth tends to decline or plateau. This is much easier to see with the adult abalone data, as the adult data tends to have less variability. The one piece of data that I find the most confusing is the 12th ring on the infant data for both Whole Weight and Volume. That box plot has the most variablility for data. Is this due to a lower sample size that is skewing the data or some other factor?    ***

**Conclusions**

Please respond to each of the following questions (8 points total):

**Question 1) (5 points)	What are plausible reasons that explain the failure of the original study? Consider to what extent abalone physical measurements may be used for predicting age.**

***Answer: I truely believe that there are many conclusions that can be made to explain the failure of the original study. Our data showed that there is extreme variability in certain sets of the data that tend to skew the results. For example, the Infant sex. The Infants not always have a uniform volume or weight based on class. This was one factor that made it extrememly difficult to write a model. I also question the other factors that were not discussed. What time of year was the study conducted? Was it before or after a recent commercial harvest? What was the temperature of the water? All of these factors could have serious consequences to collecting the right sample for a model. I think that the harvest factor could be the biggest factor to altering the data.***

**Question 2) (3 points) Setting the abalone data and analysis aside, if you were presented with an overall histogram and summary statistics from a sample and no other information, what questions might you ask before accepting them as representative of the sampled population?**

***Answer: There are a lot of factors that would come to mind if I was given a histogram and summary statistics. Some of the questions I would ask would include: How big is the population?, Was the sample obtained through random or non-random selection? Who collected the sample? Where did the sample come from? Are all locations representative of the location you choose for the sample? These questions answer alot of the important topics that an ethical and accurate analysis must entail. We must must have ethical collection of data by non-biased sources and have a thorough understanding of the population.   ***

**Question 3)  (2 points)	What do you see as difficulties when drawing conclusions from observational studies? Can causality be determined?  What might be learned from such studies?**

***Answer: Observational studies are unique and have their own benefits and challenges. Primarily, in an observational study a researcher has no control over the variables and this was the case with the abalone. There were certain limitations and variables couldn't be tweaked to test different hypothesis. A researcher couldn't make an abalone get more rings, for example. In this situation, we didn't have enough of the right data to determine causality. I think you would need to check some of the other variables I suggested in the question above. For example, You could measure abalone growth vs temperature of water. I think that observational studies have a unique purpose. They do a good job of giving you a base line for a study. They can generate new questions, which in turn can help you develop a new hypothesis to better test the same study or start a new one.    ***

